---
title: "Spotify"
Author: "Donald VM, @goldblatchickens" 
output: html_notebook
---
#Setups

```{r libs}
library(tidyverse)
```

```{r data import}
data1 <- tidytuesdayR::tt_load(2020, week = 4)
data <- data1$spotify_songs

```

# Kanye West
```{r plots for Kanye}

# Select the Kanye music
p1 <- data %>% select(track_artist,track_name, track_album_name,track_album_release_date,track_popularity,energy, danceability, instrumentalness,liveness,speechiness,tempo,duration_ms) %>% filter(track_artist == "Kanye West") %>% group_by(track_album_name) %>% mutate(average_pop = mean(track_popularity), Type_of_Kanye =
         ifelse(track_album_release_date > 2013-01-01, "New Kanye", "Old Kanye"))

p1 # a check

# making the Plot
kanye1<- p1 %>%  ggplot(mapping = aes(x = average_pop, y = danceability, color = Type_of_Kanye, label =track_album_name))+
  geom_jitter()+
  labs(x = "Album popularity", y = "Danceability", title = "Kanye West Danceability vs Album Popularity", subtitle = "Albums after MBDF are new Kanye")+
  theme_classic()


# Add the album names to the Dots
kanye1 <- kanye1+ geom_text(check_overlap = TRUE, nudge_x = 0.52, nudge_y = 0.02)

kanye1

# A second way to add the album names to the plots

library(ggrepel) # using this other library
kanye1+geom_label_repel(data = subset(p1, danceability > .55),                   box.padding   = 0.2, 
                  point.padding = 0.1,
                  segment.color = 'grey50')

# Another plot Time VS Danceability

kanye2 <- p1 %>%  ggplot(mapping = aes(x =track_album_release_date , y =danceability , color = Type_of_Kanye, label =track_album_name))+
  geom_jitter()+
  labs(x = "Release Date", y = "Danceability", title = "Kanye West vs Time VS Danceability", subtitle = "Albums after MBDF are new Kanye")+
  theme_classic()

kanye2<- kanye2+ geom_text(check_overlap = TRUE, nudge_x = 0, nudge_y = 0)

kanye2+ theme(axis.text.x = element_text(angle = 90))


# Antoher Plot 
kanye3 <- p1 %>%  ggplot(mapping = aes(x =track_album_release_date , y = speechiness , color = Type_of_Kanye, label = track_popularity))+
  geom_jitter()+
  labs(x = "Release Date", y = "speechiness", title = "Kanye West vs Time VS speechiness", subtitle = "Albums after MBDF are new Kanye")+
  theme_classic()
kanye3+ theme(axis.text.x = element_text(angle = 90))+ geom_text(check_overlap = TRUE, nudge_x = 0, nudge_y = 0)



```
## some notes on code
By adding '+ theme(axis.text.x = element_text(angle = 90))' to my ggplot, My Release Date labels (on the x-axis) are vertical and thus readable. 


# Rap & Pop music plots 

Super inspired by 

https://github.com/iandouglasbell/Tidy-Tuesdays/blob/master/Jan%2021%202020/Spotify.R

```{r 2000-2020}
library(lubridate)

# What distinc genres are there? 
unique(data$playlist_genre)

# prep data for plot
annual_trends <- spotify_songs2 %>%
  distinct(track_name, track_artist, .keep_all = TRUE) %>%
  mutate(year_released=ymd(track_album_release_date, truncated = 2L)) %>%
  mutate(year_released = year(year_released)) %>%
  filter(year_released >= 2000, playlist_genre == "pop" | playlist_genre == "rap" ) %>% mutate(duration_min = duration_ms/60000) %>%
  group_by(year_released) %>%
  summarise(Danceability = mean(danceability), Energy = mean(energy) , "Loudness (db)" = mean(loudness), "Happiness (Valence)" = mean(valence), "Tempo (BPM)" = mean(tempo), "Duration (Min)"= mean(duration_min))%>% pivot_longer(-year_released, names_to = "Parameter", values_to = "score")

annual_trends # Look at this new data. 
pivot_longer() # names_to and values_to are creating new colnames that summarise the data in a pivot. for more info check out ?pivot

# Plot it
ggplot(annual_trends) + 
  geom_point(aes(x=year_released, y=score), size=2, color="darkorange3") +
  geom_smooth(aes(x=year_released, y=score), color="grey64", se=FALSE, size=1.4) + 
  facet_wrap(~Parameter, scales = "free") + theme_minimal() + 
theme(text = element_text(size=14, color = "darkorange3")) + 
labs(title = "RAP & POP Trends 2000-2020", subtitle ='Shorter, Sadder, Quieter. A Dance Comeback?' ,caption = 'Source: Spotify | #TidyTuesday #R4DS', y="", x="") +
  theme(plot.background = element_rect(fill = "pink1"))+
  theme(plot.title = element_text(size=18, face="bold"))

# save this plot for people that need the right type of image. 
ggsave("pop_rap.png", dpi = 300, width = 12, height = 9, units = "in")

```


```{r oldies}
annual_trends2 <- spotify_songs2 %>%
  distinct(track_name, track_artist, .keep_all = TRUE) %>%
  mutate(year_released=ymd(track_album_release_date, truncated = 2L)) %>%
  mutate(year_released = year(year_released)) %>%
  filter(year_released %in% c(1990:2000)) %>% mutate(duration_min = duration_ms/60000) %>%
  group_by(year_released) %>%
  summarise(Danceability = mean(danceability), Energy = mean(energy) , "Loudness (db)" = mean(loudness), "Happiness (Valence)" = mean(valence), "Tempo (BPM)" = mean(tempo), "Duration (Min)"= mean(duration_min))%>% pivot_longer(-year_released, names_to = "Parameter", values_to = "score")

annual_trends2 # Look at this new data. 
#pivot_longer() => names_to and values_to are creating new colnames that summarise the data in a pivot. for more info check out ?pivot

# Plot it
ggplot(annual_trends2) + 
  geom_point(aes(x=year_released, y=score), size=2, color="darkorange3") +
  geom_smooth(aes(x=year_released, y=score), color="grey64", se=FALSE, size=1.4) + 
  facet_wrap(~Parameter, scales = "free") + theme_minimal() + 
theme(text = element_text(size=14, color = "darkorange3")) + 
labs(title = "The 90's", subtitle ='Fuck yhea' ,caption = 'Source: Spotify | #TidyTuesday #R4DS', y="", x="") +
  theme(plot.background = element_rect(fill = "pink1"))+
  theme(plot.title = element_text(size=18, face="bold"))

# save this plot for people that need the right type of image. 
ggsave("the_90s.png", dpi = 300, width = 12, height = 9, units = "in")
```

# Using coordinate Polar and Radar

**TL;DR This is a test and it didn't really do what I wanted.** 

Thanks to mike for the inspiration, although what he did is far far cooler
https://github.com/mikemahoney218/tidytuesday-projects/blob/master/2020/2020-01-21/generator.Rmd

```{r making a polar plot}
 
data %>% select(valence, speechiness,danceability,energy, playlist_genre) %>% filter(playlist_genre=="pop") %>% summarise_all(mean) %>% select(-playlist_genre) %>% gather(dimension, val) %>% ggplot(aes(dimension, val, fill = dimension)) +
    geom_col(color = "black") +
    scale_fill_brewer(name = "",
                      palette = "Accent") +
    coord_polar() +
    theme_minimal()+
    labs(x = "",
         y = "",
         title = "Pop's average song characteristics")+
  theme(plot.background = element_rect(fill = "grey100"))+
  theme(plot.title = element_text(size=11, face="bold"))+
  theme(legend.position = "none")
  

"pink1"


# This isn't what I want, I would love to have a facet wrap but it didnt work. 
data %>% select(valence, speechiness,danceability,energy, playlist_genre) %>% filter(playlist_genre=="pop") %>% summarise_all(mean)%>% gather(dimension, val,-playlist_genre) %>% ggplot(aes(dimension, val, fill = dimension)) +
    geom_col(color = "black") +
    scale_fill_brewer(name = "",
                      palette = "Accent")+
    coord_polar() +
    theme_minimal()+
    labs(x = "",
         y = "",
         title = "Pop's average song characteristics")+
  theme(plot.background = element_rect(fill = "grey100"))+
  theme(plot.title = element_text(size=11, face="bold"))+
  theme(legend.position = "none")



# Radar plot 

library(fmsb)
data %>% select(valence, speechiness,danceability,energy, playlist_genre) %>% group_by(playlist_genre) %>% summarise_all(mean)%>% gather(dimension, val,-playlist_genre) %>% ggplot(aes(dimension, val, fill = dimension)) +
    geom_col(color = "black") +
    scale_fill_brewer(name = "",
                      palette = "Accent")+
    radarchart() +
    theme_minimal()+
    labs(x = "",
         y = "",
         title = "Pop's average song characteristics")+
  theme(plot.background = element_rect(fill = "grey100"))+
  theme(plot.title = element_text(size=11, face="bold"))+
  theme(legend.position = "none")


```

Wooow What A struggle this is. 

some help over at: https://gitlab.com/choh/tidy-tuesday/blob/master/R/2020_04_music_genres.R

but im tired and not very exited about this anymore. 

This guy did teach me something I can use for my kanye plots 
songs <- data %>% distinct(track_id, .keep_all = TRUE)

